### 2019-09-08 ic-journal/response-letter.tex

\vspace{1cm}

\section*{Responses to Review 1 (accept - minor revision)}

\subsection*{Meta comments}

\begin{enumerate}
\item \begin{quote}
    This paper describes a HOL formalisation of several notions in
    process theory: CCS and various kinds of bisimilarity, and in
    particular the unique-solutions technique. The latter was
    introduced by Milner, and recently revisited and refined by the
    second author.
  \end{quote}

\item \begin{quote}
    In general, these kinds of results are error-prone, especially
    those related to weak bisimilarity: there are multiple known
    mistakes in the literature. Therefore, formalisation is a
    worthwile effort. Moreover, the formalisation itself is quite
    readable. The paper itself documents this formalisation and is,
    for the most part, clear.
  \end{quote}

\item \begin{quote}
    The paper is an extended version of an EXPRESS/SOS paper. The
    extension consists of a more detailed report of the proofs -
    perhaps this is enough.
  \end{quote}

@@

\item \begin{quote}
    In conclusion, on the positive side, I think the authors did some
    important work formalising part of the theory of CCS in HOL, and
    on the negative side I have doubts about the scientific value of
    the article they wrote about it. I am therefore reluctant to
    support publication in Information and Computation at this stage.
  \end{quote}

\item \begin{quote}
    I am therefore reluctant to support publication in Information and
    Computation at this stage. I suggest to ask the authors for a
    major revision in which they make the paper accessible for readers
    not familiar with HOL, comment in more detail on the HOL proof
    scripts and what can be learned from them, and take into account
    the detailed comments below.
  \end{quote}
\end{enumerate}

\subsection*{Line-based comments}

\begin{enumerate}
\item \begin{quote}
    Abstract:
    There should be a remark that this is about weak bisimilarity. In
    general 'bisimilarity' is strong bisimilarity.
  \end{quote}
  We have added "(weak)" before "bisimilarity", in the first line of abstract.

\item \begin{quote}
    L18: Remove the word 'rather'
  \end{quote}
  Removed.

\item \begin{quote}
    L28: Hennessy Lemma -> Hennessy's Lemma
  \end{quote}
  Done.

\item \begin{quote}
    L28: Deng Lemma -> Deng's lemma
  \end{quote}
  Done.

\item \begin{quote}
    L28: Remove the word 'long'
  \end{quote}
  Done.

\item \begin{quote}
    L28: proofs -> theorems
  \end{quote}
  Done.

\item \begin{quote}
    L32: Add the word 'also' after 'the work is'
  \end{quote}
  Done.

\item \begin{quote}
    L34: From the view of -> From the point of view of
  \end{quote}
  Done.

@@

\item \begin{quote}
    L35: Regarding the sentence "as formally proving a previously known result gives us the chance to see
    what is really needed for establishing that result."
    Either omit this sentence or explain what technical insights one
    gains by doing this work
  \end{quote}

  Our formal proof of the classical result of ``$\rapprox$ is the
  coarsest congruence contained in $\wb$'' is one such example, where
  the formal theorem we obtained has slightly weaker antecedents under
  the same proof idea. Anyway, we have omitted the sentence here.

@@

\item \begin{quote}
    L59: "shown in Fig. 1" is placed after "a Labeled Transition
    System", causing the belief that Figure 1 contains an LTS. 
    But there is no LTS. Figure 1 contains the SOS of CCS.
  \end{quote}

  We have rephrased it, saying Fig.1 represents an LTS expressed in SOS rules.

\item \begin{quote}
    Semantics: The relabeling rule is not explained anywhere.
  \end{quote}

\item \begin{quote}
    Semantics: The constraint on the relabeling functions does not belong to the inference rule, but should be formulated as a restriction on the syntax.
  \end{quote}
  We have moved the side condition to the main text where we explain
  the syntax. (TODO)

\item \begin{quote}
    L74: Add the word 'a' between 'are' and 'context'.
  \end{quote}
  Done. (``... there exists a context ...'')

@@

\item \begin{quote}
    L74: Add the word 'a' between 'are' and 'context'.
  \end{quote}
  Done. (``... there exists a context ...'')

\item \begin{quote}
  Contexts: Does an $i$-holed context have exactly one hole $[]_i$ for
  each $i$? This question is not answered anywhere in the text.

  In the formalisation you only use contexts with one kind of hole $[]$. This definition can be changed to only
  use one hole as well, with a clear indication that this hole can occur multiple times.
\end{quote}
TODO

\item \begin{quote}
    L82: Remove the comma after $Q'$
  \end{quote}
  Done.

\item \begin{quote}
    Definition 2.2: Superscript $c$ is already used for contextual closure.
    Does $\approx^c$ stand for the contextual closure of $\approx$, or is it something else?
    Since contextual closure is not used anywhere in the paper, it is best to remove it.
    Otherwise you need different notation for rooted bisimilarity.
  \end{quote}
  TODO

\item \begin{quote}
    L91: What does 'reducing' mean?
  \end{quote}
We have rephrased the sentence: ``The above definition brings up a
proof technique for proving rooted bisimilarity results, starting from
a given bisimulation.''

\item \begin{quote}
    L91: bisimiarity -> bisimilarity
  \end{quote}
  Fixed.

\item \begin{quote}
    L92: bisimiarity -> bisimilarity
  \end{quote}
  Fixed.

@@

\item \begin{quote}
    L92: Add a colon after 'results'
  \end{quote}

\item \begin{quote}
    L118: Why is the word 'actions' in bold font? It says 'ds' in the margin, what does that mean?
  \end{quote}
  We have corrected it. It's an internal comment.

\item \begin{quote}
Definition 3.3: The definition of ‘sequential’ is awkward.  Moreover
this definition calls for some examples since the definitions can be
very confusing.
\end{quote}
We have added some examples, and used Milner's original definition of sequentiality.

\item \begin{quote}
Theorem 3.4: What are direct sums?
\end{quote}
Rephrased: ``with guarded sums only''. We have removed all usages of
``direct sym''.

\item \begin{quote}
    L.127: I think it would be clearer to give two concrete solutions
    rather than saying that any process not using a is a solution.
  \end{quote}
  Done. Two examples are $\nil$ and $b.\nil$.

\item \begin{quote}
    L138: "Bisimilarity contraction ... some contraction R". This is not a grammatically correct sentence.
  \end{quote}
  We have rephrased the sentence.

\item \begin{quote}
    L142: ".. are even required". The word 'even' suggests that this
    is a stronger requirement, but it is in fact weaker.
  \end{quote}
  We have replaced it with ``indeed''.

\item \begin{quote}
    L143: Add the words "that is" after 'preorder'.
  \end{quote}
  Done.

\item \begin{quote}
    L144: remove ‘that’
  \end{quote}
  Done.

@@

\item \begin{quote}
    L153: "As bisimilarity .. but sums." This is not a grammatically correct sentence.
  \end{quote}

  We have removed the sentence, including this statement with the
  statement after the example.

@@

\item \begin{quote}
    Example 3.7: this example shows strictness of the inclusion. But
    the sentence before the example suggests that this is going to
    show that the preorder is not preserved by sums.
  \end{quote}

  Having removed the sentence above now this should be ok. We have
  also rephrased the sentence following the Example 3.7.

\item \begin{quote}
    L157: Doesn’t this statement deserve a proof?
  \end{quote}
  We have added a reference which contains this proof. Additionally,
  this proof will be discussed later in a formalized version.

@@

  It should be ``via''.

\item \begin{quote}
    L184: closes -> completes
  \end{quote}
  Done.

\item \begin{quote}
    L194: ‘the definition should not be recursive’?
  \end{quote}

\item \begin{quote}
    Definition 3.11: Similar to the remark about Definition 2.2.
  \end{quote}
  TODO

\item \begin{quote}
    Proof of Theorem 3.14: Why do you use context $C$ instead of tuples of expressions $\tilde{E}$?
    If this is the case because you only use contexts with one kind of hole, then you should use the letter
    E instead of C.
    If $\tilde{P}$ is a solution then $P_i >=^c E_i[\tilde{P}]$ for all i, but this quantification over i is missing.
  \end{quote}
  We have completely re-worked this proof.

\item \begin{quote}
    L215: The prove -> The proof
  \end{quote}
  Fixed during re-work of the proof of Theorem 3.14.

\item \begin{quote}
    General: the word 'Logic' in "HOL Logic" is redundant.
  \end{quote}
  Removed ``Logic'' from ``HOL Logic''.

\item \begin{quote}
    L223-225: Perhaps you want to reformulate this long sentence. It is hard to read.
  \end{quote}
  The long sentence is spilited.

\item \begin{quote}
    L227: considerable -> considerably
  \end{quote}
  Done.

\item \begin{quote}
    L228: is -> are
  \end{quote}
  Done.

\item \begin{quote}
    L228: what does convincible mean here?
  \end{quote}
  TODO

\item \begin{quote}
    L230: single?
  \end{quote}
  Changed to: ``the \emph{same} (programming language)''

\item \begin{quote}
    L235: The sentence between parentheses is an important remark and
    should not be between parentheses.
  \end{quote}
  Removed parentheses.

\item \begin{quote}
    L238: without loss of generality?
  \end{quote}
  We have eliminated this sentence as now we have done the
  multivariate formalizations.

\item \begin{quote}
    L239: similar to the comment about line 118, theorem-prover is
    bold and 'ds' is in the margin.
  \end{quote}
  We have eliminated this sentence as now we have done the
  multivariate formalizations.

\item \begin{quote}
    L245: now based on HOL’s option theory?
  \end{quote}
  \texttt{option} is the name of the HOL formal theory.

\item \begin{quote}
    L249: What are ‘backquotes’?
  \end{quote}
  We have eliminated most uses of backquotes using HOL's new syntax.

\item \begin{quote}
    Footnote 3: What are CCS literals?
  \end{quote}
  Sorry, it should be ``literature''.

\item \begin{quote}
    L265: What does the turnstile mean? I would expect that structural
    operational rules are axioms of the theory, but the turnstile
    suggests to me that they are theorems.
  \end{quote}
  TODO

\item \begin{quote}
    L275: we substitute rec A. P for all occurrences of the variable A in P.
  \end{quote}
  Done.

\item \begin{quote}
    L279: substutiion -> substitution
  \end{quote}
  Fixed.

\item \begin{quote}
    L289: these -> the, on the tenary -> of the ternary
  \end{quote}
  Fixed.

\item \begin{quote}
    General: in the lines 293-300 the trailing deadlock 0 is given,
    while in the preliminaries you stated that it would be omitted. If
    you decide to keep the deadlock, it should also be in bold font.
  \end{quote}
  We have made them all in bold fonts.

\item \begin{quote}
    L296: I don’t understand fact (1).
  \end{quote}
We have re-phrased this part: ``(1) there exist indeed three
  transitions mentioned above; (2) there exists no other
  transition.''

\item \begin{quote}
    L299: ‘Hence ... CCS_TRANS_CONV’ is not a grammatically correct sentence; there is a verb missing.
  \end{quote}
  We have rephrased this sentence.

\item \begin{quote}
    L316: Explain what is p.
  \end{quote}
  TODO

\item \begin{quote}
    Footnote 4: Do you mean ‘unguarded recursion’? How can relabelling operators cause infinite branching?
  \end{quote}
  We have removed the part ``due to the use of ...''. Infinite
  branching is indeed possible, no matter what's the reason.

\item \begin{quote}
    L362: tedious but long -> tedious and long
  \end{quote}
  Done.

\item \begin{quote}
    L374: An highlight -> A highlight
  \end{quote}
  Rephrased to ``One highlight...''

\item \begin{quote}
    L378: relations on CCS process -> relation on CCS processes
  \end{quote}
  Done.

\item \begin{quote}
    Footnote 5: Add the word 'respectively' between 'quantifiers' and 'in'.
  \end{quote}
  Done.

\item \begin{quote}
    Footnote 6: Regarding the last sentence between parentheses. What do you mean by this?
  \end{quote}
  By default, all theorems (their \TeX code) exported from HOL are
  fully specialised, removing outermost universal quantifiers. We have
  moved this sentence to an earlier footnote on the first such exported theorem.

\item \begin{quote}
    General: The same trick of applying the coinduction package to the
    definition of bisimulation is repeated twice.
  \end{quote}
  TODO

\item \begin{quote}
    General: In the sections about the HOL formalisation you refer to
    weak bisimulation instead of bisimulation.    
  \end{quote}
  TODO

\item \begin{quote}
    L461: I think there is no need to include those 3
    theorems. Wouldn’t it suffice to write that similar theorems are
    returned as for \texttt{STRONG\_EQUIV}?
  \end{quote}
  TODO

\item \begin{quote}
    L496-503: A similar remark can be made for \texttt{STRONG\_EQUIV}. Why do you put it here?
  \end{quote}
  TODO

\item \begin{quote}
    L523,524: Similar to the comment about line L28
  \end{quote}
  Done.

\item \begin{quote}
    L545: What does this statement mean?
  \end{quote}
  We have rephrased this sentence. Basically we meant, for whatever
  algebraic law of strong bisimilarity, there're also versions for
  weak and rooted bisimilarity.

@@

\item \begin{quote}
    L545: What does this statement mean?
  \end{quote}
  We have rephrased this sentence. Basically we meant, for whatever
  algebraic law of strong bisimilarity, there're also versions for
  weak and rooted bisimilarity.

\item \begin{quote}
    Up-to techniques: You mention four up-to techniques for weak
    bisimilarity and each of them are not powerful enough to prove the
    unique solution theorems. Why are they not powerful enough?
  \end{quote}

  We have repharsed this part, moving the citation
  \cite{sangiorgi1992problem} (which explains the reason) right before
  the first sentence.

\item \begin{quote}
    L580: ‘Surprisingly’ To me this is not really surprising. Why do you find it surprising?
  \end{quote}
  Actually this is indeed surprising. We have rephrased this sentence:
  ``this is mostly due to the surprising fact
  that, although the contraction preorder is contained in weak
bisimilarity, contraction does not imply weak bisimulation (while
expansion does so)''.

\item \begin{quote}
    L606-619: I really see no added value in repeating Def. 4.2 in HOL syntax.
  \end{quote}
  We have removed the HOL version.

\item \begin{quote}
    Def. 4.4: Shouldn’t there be double arrows? (twice)
  \end{quote}
  Yes. We have fixed the definition.

\item \begin{quote}
    L700: This paragraph starts with a sentence in parentheses, which is not very desirable.
    Moreover this sentence does not seem to be connected to the next sentence.
  \end{quote}
  We have rephrased the sentence.

\item \begin{quote}
    L701: Sentence does not end with a period.
  \end{quote}
  We have added a colon at the end.

\item \begin{quote}
    L704: What do you mean by "which itself may not be"?
  \end{quote}
  We have rephrased the sentence: ``(which needs not to be itself a (pre)congruence)''.

\item \begin{quote}
    Figure 2: The $\subseteq$ from \texttt{SUM\_EQUIV} to Rooted
    bisimilarity makes the figure confusing; put a $\supseteq$
  \end{quote}
  Done.

\item \begin{quote}
    L723: is -> in
  \end{quote}
  Done.

\item \begin{quote}
    L723-L727: This sentence should be split into multiple sentences.
  \end{quote}
  We have rephrased the sentence.

\item \begin{quote}
    L744-745: pure mathematics?
  \end{quote}
  We have removed this sentence.

\item \begin{quote}
    L746: Arbitrary -> Arbitrarily
  \end{quote}
  Done.

\item \begin{quote}
    L749: "infinite such sub-processes" this construction does not seem right and could be avoided.
  \end{quote}
  We have rephrased the sentence: ``There could be infinitely many such sub-processes, even on finitely
  branching processes.''

@@

\item \begin{quote}
    L754: non-$\tau$ action $a$ -> $a\neq\tau$
  \end{quote}

\item \begin{quote}
    L766: "By induction on the definition", don't you mean by induction on natural numbers?
  \end{quote}
  We have rephrased the sentence: ``Following the inductive structure
  of the definition of Klop processes,...''

\item \begin{quote}
    L823: are -> is
  \end{quote}
  Done.

\item \begin{quote}
    L831: `In above proof’ Which proof are your referring too? I only see a theorem.
  \end{quote}
  Fixed as: ``In the proof of the above theorem''.

\item \begin{quote}
    L837: Add the word 'a' between 'not' and 'congruence'
  \end{quote}
  Done.

\item \begin{quote}
    L837: why do you also need the condition of sequentiality?
  \end{quote}
  TODO.

\item \begin{quote}
    L991: error-prune -> error-prone, ’since one is tempted to overlook details’?
  \end{quote}
  We have rephrased the sentence.

\item \begin{quote}
    L1001: the trace equivalences -> trace-based equivalences
  \end{quote}
  Done.

\item \begin{quote}
    L1006: in alphabetic order
  \end{quote}
  Done.

\item \begin{quote}
    L1040: hol4 -> HOL4
  \end{quote}
  Fixed bibtex item.

### 2019-09-08 ic-journal/ccs.tex
Stage fixes with highlights
Merged all revise suggestions from Sangiorgi

We assume a possibly infinite set of names $\mathscr{L} = \{a, b,
\ldots\}$ forming input and $\overline{\mbox{output}}$ actions, plus a special invisible
action $\tau \notin \mathscr{L}$, and a set of variables $A, B,
\ldots$ for defining recursive behaviours.
Given a deadlock $\nil$, the class of CCS processes is inductively defined from $\nil$ by the operators
of prefixing, parallel composition, summation (binary choice), restriction, recursion and relabeling:
\begin{equation*}
\begin{array}{cccl}
\mu  & := & \tau\!\!\!\! & \midd \; a  \; \midd \;  \outC a  \\
P  & := & \nil\!\!\!\! & \midd \;  \mu . P \; \midd \;  P_1 |  P_2 \; \midd  \;
P_1 + P_2 \; \midd % \; \mu . P\; \midd  \; 
  (\res a\!)\, P  \;  \midd \;  A \; \midd \; \recu A  P
\; \midd \; P\; [r\!f]  % relabelling
\end{array}
\end{equation*}
We sometimes omit the trailing $\nil$, e.g., writing $a|b$ for $a.\nil |b .\nil$.
The operational semantics of CCS is then given by means of
a Labeled Transition System (LTS) \hl{expressed in Structural Operational
Semantics (SOS) rules}, shown in Fig.~\ref{f:LTSCCS}.
The symmetric version of the two \hl{SOS} rules for
parallel composition and the rule for sum are omitted.
A CCS expression uses only \emph{weakly guarded sums} if all occurrences of
the sum operator are of the form $\mu_1.P_1 + \mu_2.P_2 + \ldots
+ \mu_n.P_n$, for some $n \geq 2$.
 The \emph{immediate derivatives} of a
process $P$ are the elements of the set $\{P' \st P \arr\mu P' \mbox{
  for some $\mu$}\}$.
We use $\ell$ to range over
visible actions (i.e.~inputs or outputs, excluding  $\tau$)
and $\alpha, \mu$  to range over all actions.
\begin{figure*}[t]
\begin{center}
\vskip .1cm
 $\displaystyle{  \over  \mu.  P    \arr\mu
P } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P + Q   \arr\mu
P'  } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P | Q   \arr\mu
P' | Q } $  $ \hb$   
\hskip .3cm
  $\; \;$  $\displaystyle{ P \arr{ a}P' \hk \hk  Q
\arr{\outC a }Q'  \over     P|  Q \arr{ \tau} P'
|  Q'  }$ 
\\
\vspace{.2cm}
$\displaystyle{ P \arr{\mu}P' \over
 (\res a\!)\, P   \arr{\mu} (\res a\!)\, P'} $ $ \mu \neq a, \outC a$
$ \hb$
%
$\displaystyle{ P \sub {\recu A P} A \arr{\mu}P' \over
 \recu A P   \arr{ \mu} P'  } $
\hskip .5cm  
$\displaystyle{ P \arr{\mu} P' \over
 P \;[r\!f] \arr{r\!f(\mu)} P' \;[r\!f]} $ $\forall a.\, r\!f(\outC a) = \overline{r\!f(a)}$
$ \hb$ %  &
\end{center}
\caption{Structural Operational Semantics of CCS}
\label{f:LTSCCS}
\end{figure*}
Some standard notations for transitions: $\Arr\epsilon$ is the 
reflexive and transitive closure of $\arr\tau$, and 
$\Arr \mu $ is $\Arr\epsilon \arr\mu \Arr\epsilon$ (the
composition of the three relations).
Moreover,   
$ P \arcap \mu P'$ holds if $P \arr\mu P'$ or ($\mu =\tau$ and
$P=P'$); similarly 
$ P \Arcap \mu P'$ holds if $P \Arr\mu P'$ or ($\mu =\tau$ and
$P=P'$).
We write $P \:(\arr\mu)^n P'$ if $P$ can become $P'$ after performing
$n$ $\mu$-transitions. Finally, $P \arr\mu$ holds if there is $P'$
with $P \arr\mu P'$, and similarly for other forms of transitions.

\paragraph{Further notations}
We let $\R$, $\S$ range over binary relations, sometimes using 
 infix notation for them; e.g., 
$P \RR Q$ means that $(P,Q) \in \R$.
We use a tilde, as in  $\til P$, to denote tuples of processes with
countably many elements (thus the tuple may also be infinite.)
All notations are extended to tuples componentwise,
e.g., $\til P \RR \til Q$ means that $P_i \RR Q_i$, for each  
component $i$ of the tuples $\til P$ and $\til Q$.
And $\ct{\til P}$ is the process obtained by replacing each hole
$\holei i$ of the  context $\qct$ with $P_i$.  
We write $
\ctx \R$ for the closure of a relation under contexts. Thus $P\: \ctx \R\: Q$
means that there are context $\qct$ and tuples $\til P,\til Q$ with
$P =  \ct{\til P}, Q =  \ct{\til Q}$ and $\til P \RR \til Q$.
We use the symbol $\DSdefi$ for abbreviations. For instance, $P \DSdefi G $, where
$G$ is some expression, means that $P$ stands for the expression $G$.
If $\leq$ is a preorder, then  $\geq$  is its inverse (and
conversely).

% next file: expansion.tex

>>>>

We assume a possibly infinite set of names $\mathscr{L} = \{a, b,
\ldots\}$ forming input and $\overline{\mbox{output}}$ actions, plus a special invisible
action $\tau \notin \mathscr{L}$, and a set of variables $A, B,
\ldots$ for defining recursive behaviours.
The class of CCS processes is inductively
defined from $\nil$ \hl{(deadlock)} by the operators
of prefixing, parallel composition, summation (binary choice),
restriction, relabeling and recursion:
\begin{equation*}
\begin{array}{cccl}
\mu  & := & \tau\!\!\!\! & \midd \; a  \; \midd \;  \outC a  \\
P  & := & \nil\!\!\!\! & \midd \;  \mu . P \; \midd \;  P_1 |  P_2 \; \midd  \;
P_1 + P_2 \; \midd 
(\res a\!)\, P
\; \midd \; P\; [r\!f]  % relabelling
\;  \midd \;  A \; \midd \; \recu A  P
\end{array}
\end{equation*}
We sometimes omit a trailing $\nil$, e.g., writing $a|b$ for $a.\nil |b .\nil$.
The operational semantics of CCS is then given by means of
a Labeled Transition System (LTS) \hl{expressed in Structural Operational
Semantics (SOS) rules}, shown in Fig.~\ref{f:LTSCCS}.
The symmetric version of the two \hl{SOS} rules for
parallel composition and the rule for sum are omitted.
A CCS expression uses only \emph{weakly guarded sums} if all occurrences of
the sum operator are of the form $\mu_1.P_1 + \mu_2.P_2 + \ldots
+ \mu_n.P_n$, for some $n \geq 2$.
 The \emph{immediate derivatives} of a
process $P$ are the elements of the set $\{P' \st P \arr\mu P' \mbox{
  for some $\mu$}\}$.
We use $\ell$ to range over
visible actions (i.e.~inputs or outputs, excluding  $\tau$)
and $\alpha, \mu$  to range over all actions.
\begin{figure*}[t]
\begin{center}
\vskip .1cm
 $\displaystyle{  \over  \mu.  P    \arr\mu
P } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P + Q   \arr\mu
P'  } $  $ \hb$   
\hskip .5cm
 $\displaystyle{   P \arr\mu   P' \over   P | Q   \arr\mu
P' | Q } $  $ \hb$   
\hskip .3cm
  $\; \;$  $\displaystyle{ P \arr{ a}P' \hk \hk  Q
\arr{\outC a }Q'  \over     P|  Q \arr{ \tau} P'
|  Q'  }$ 
\\
\vspace{.2cm}
$\displaystyle{ P \arr{\mu}P' \over
 (\res a\!)\, P   \arr{\mu} (\res a\!)\, P'} $ $ \mu \neq a, \outC a$
$ \hb$
%
$\displaystyle{ P \sub {\recu A P} A \arr{\mu}P' \over
 \recu A P   \arr{ \mu} P'  } $
\hskip .5cm  
$\displaystyle{ P \arr{\mu} P' \over
 P \;[r\!f] \arr{r\!f(\mu)} P' \;[r\!f]} $
$ \hb$ %  &
\end{center}
\caption{Structural Operational Semantics of CCS}
\label{f:LTSCCS}
\end{figure*}
Some standard notations for transitions: $\Arr\epsilon$ is the 
reflexive and transitive closure of $\arr\tau$, and 
$\Arr \mu $ is $\Arr\epsilon \arr\mu \Arr\epsilon$ (the
composition of the three relations).
Moreover,   
$ P \arcap \mu P'$ holds if $P \arr\mu P'$ or ($\mu =\tau$ and
$P=P'$); similarly 
$ P \Arcap \mu P'$ holds if $P \Arr\mu P'$ or ($\mu =\tau$ and
$P=P'$).
We write $P \:(\arr\mu)^n P'$ if $P$ can become $P'$ after performing
$n$ $\mu$-transitions. Finally, $P \arr\mu$ holds if there is $P'$
with $P \arr\mu P'$, and similarly for other forms of transitions.

\paragraph{Further notations}
We let $\R$, $\S$ range over binary relations, sometimes using 
 infix notation for them; e.g., 
$P \RR Q$ means that $(P,Q) \in \R$.
We use a tilde, as in  $\til P$, to denote tuples of processes with
countably many elements (thus the tuple may also be infinite.)
All notations are extended to tuples componentwise,
e.g., $\til P \RR \til Q$ means that $P_i \RR Q_i$, for each  
component $i$ of the tuples $\til P$ and $\til Q$.
And $\ct{\til P}$ is the process obtained by replacing each hole
$\holei i$ of the  context $\qct$ with $P_i$.  
We write $\ctx \R$ for the closure of a relation under contexts.
Thus $P\: \ctx \R\: Q$
means that there \hl{exists a} context $\qct$ and tuples $\til P,\til Q$ with
$P =  \ct{\til P}, Q =  \ct{\til Q}$ and $\til P \RR \til Q$.
We use the symbol $\DSdefi$ for abbreviations. For instance, $P \DSdefi G $, where
$G$ is some expression, means that $P$ stands for the expression $G$.
If $\leq$ is a preorder, then  $\geq$  is its inverse (and
conversely).

% next file: expansion.tex (bisimilarity and rooted bisimilarity, actually)
